Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-25 paper.qmd
Typo fix
More on Parkin approach

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography:
  # cruse.bib
  - bibliography.bib
  - cruse.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  # html: default
  # docx: default
# # # # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # # #   # pdf: default

@@

This describes describe the design, features and potential use of an approach to strategic cycle network planning that is open, scalable, and evidence-based.
We label the approach Cycle Route Uptake and Scenario Estimation (CRUSE) and present a case study of its application in Ireland, the results of which are publicly available at [cruse.bike](https://cruse.bike/) in the CRUSE Tool for Ireland, the landing page of which is illustrated in @fig-landing.
In @sec-methods, we outline the methods used to generate the evidence presented in the CRUSE Tool.
In @sec-results, we present the results of the CRUSE Tool, including estimates of current cycling levels and future cycling potential at the national, regional and local levels.
In @sec-discussion, we discuss limitations and possible future improvements to the approach, and the implications of the results for Ireland before making concluding remarks in @sec-conclusions.

@@

Tools for estimating geographical distribution of cycling potential have advanced substantially in recent years.
Developed by researchers and developers in academic, public and private sectors, they have evolved from simple area-based static models tailored to specific regions to more complex and (in some cases) more generalizable tools.
Outputs include maps with results available at several levels of analysis with levels of availability ranging from being only available to researchers and in static maps to open access mapping systems available to the public.
Some of the most relevant existing tools are summarised below and presented in @tbl-tools.

@@

Area-based tools were developed first.
Parkin and colleagues presented a regression model to estimate the percentage of cycling trips to work in England and Wales based on socioeconomic, transportation, and physical factors at the small area (wards) level in 2007 [@parkin_estimation_2007].
They found that areas with a higher percentage of females, non-whites, car ownership, lower socioeconomic classes, income, distance to work, population density, poor highway conditions, hills, rainfall, and fewer off-road bicycle routes tended to have a lower proportion of cycling to work.

@@

>>>>

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography:
  # cruse.bib
  - bibliography.bib
  - cruse.bib
  - references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  # html: default
  # docx: default
# # # # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # # #   # pdf: default

@@

This paper describes an approach to strategic cycle network planning tool development that is open, scalable, and evidence-based.
We label the approach Cycle Route Uptake and Scenario Estimation (CRUSE) and present a case study of its application in Ireland, the results of which are publicly available at [cruse.bike](https://cruse.bike/) in the CRUSE Tool for Ireland, the landing page of which is illustrated in @fig-landing.
In @sec-methods, we outline the methods used to generate the evidence presented in the CRUSE Tool.
In @sec-results, we present the results of the CRUSE Tool, including estimates of current cycling levels and future cycling potential at the national, regional and local levels.
In @sec-discussion, we discuss limitations and possible future improvements to the approach, and the implications of the results for Ireland before making concluding remarks in @sec-conclusions.

@@

Tools for estimating geographical distribution of cycling potential have advanced substantially in recent years.
Developed by researchers and developers in academic, public and private sectors, they have evolved from simple area-based static models tailored to specific regions to more complex and (in some cases) more generalizable tools.
These tools can be considered as a specific application of the concept of planning support systems (PSS) , which are designed to help planners and decision-makers to make more informed decisions [@geertman2009].
Outputs include maps with results available at several levels of analysis with levels of availability ranging from being only available to researchers and in static maps to open access mapping systems available to the public.

Previous studies have applied the concept of PSS to specific modes, including walking [@bencekri2024], public transport [@barmentlo2012] and cycling [@bencekri2023].
While such prior work provides insight into the potential for approaches to transport planning that are both data-driven and participatory, the focus of this paper is on tools that focus attention on the *geographic distribution* of cycling potential and which answer the question "where to build".
Not all papers reviewed in this section classify themselves as, or even mention, PSS, but all of them can be considered as PSS.
This paper is informed by the thinking underlying PSS, including the idea that the resulting tools are more effective if they are modular and can inter-link with other tools and processes in the planning process: PSS can be conceived as a "toolbox with separate instruments that can work and talk to each other effectively (as in the open systems concept) and that will be selectively applied in varying configurations to support a particular phase in the planning process" [@geertman2002].
We share this conception of PSS, emphasizing the benefits of a focus on a particular mode (cycling) and a particular phase in the planning process (strategic network development).

Some of the most relevant existing tools are summarised below and presented in @tbl-tools.

@@

Area-based approaches were developed before network-based tools, such as a 2007 regression-based model to estimate the percentage of cycling trips to work in England and Wales based on socioeconomic, transportation, and physical factors at the small area (wards) level in 2007 [@parkin_estimation_2007].
They found that areas with a higher percentage of females, non-whites, car ownership, lower socioeconomic classes, income, distance to work, population density, poor highway conditions, hills, rainfall, and fewer off-road bicycle routes tended to have a lower proportion of cycling to work.
The approach also generated "forecasts for potential levels of bicycle use" with with increases in off-road routes, improvements of highway and, conversely, increases in car ownership and distance to work leading to decreases in cycling to work.

@@

CRUSE is a "great tool for getting started with project outline documents". It's an "ideal tool for generating a strategic rationale for estimating demand and building the business case in the absence of cycle count data" (Declan Keenan, Senior Transport Planner working on active travel, personal communication).
"It's a good representation of what's going on"

"I like to know that the data on the county maps is from POWSCAR data which is gold standard".

### 2024-03-25 paper.tex
More on Parkin approach

Transport systems dominated by heavy and powerful private cars are
inefficient, dangerous, and unhealthy. Cars are a major and growing
source of greenhouse gas emissions causing climate change
\citep{winkler2023}, a leading cause of premature death and injury due
to road traffic collisions \citep{globals2018} , and a cause of disease
due to physical inactivity and air, noise and (from tyres) microplastic
pollution \citep{mattsson2023} \citep{welch2023} \citep{cavallaro2024}.
Transport is responsible for 23\% of global greenhouse gas (GHG)
emissions, 70\% of which are from road transport. Nearly half of these
emissions (around 10\% of global emissions) are from passenger cars
\citep{jaramillo2022}. The transport system encourages, enables and in
some cases enforces unsustainable lifestyles. Services that are only
accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

@@

This describes describe the design, features and potential use of an
approach to strategic cycle network planning that is open, scalable, and
evidence-based. We label the approach Cycle Route Uptake and Scenario
Estimation (CRUSE) and present a case study of its application in
Ireland, the results of which are publicly available at
\href{https://cruse.bike/}{cruse.bike} in the CRUSE Tool for Ireland,
the landing page of which is illustrated in Figure~\ref{fig-landing}. In
Section~\ref{sec-methods}, we outline the methods used to generate the
evidence presented in the CRUSE Tool. In Section~\ref{sec-results}, we
present the results of the CRUSE Tool, including estimates of current
cycling levels and future cycling potential at the national, regional
and local levels. In Section~\ref{sec-discussion}, we discuss
limitations and possible future improvements to the approach, and the
implications of the results for Ireland before making concluding remarks
in Section~\ref{sec-conclusions}.

@@

Tools for estimating geographical distribution of cycling potential have
advanced substantially in recent years. Developed by researchers and
developers in academic, public and private sectors, they have evolved
from simple area-based static models tailored to specific regions to
more complex and (in some cases) more generalizable tools. Outputs
include maps with results available at several levels of analysis with
levels of availability ranging from being only available to researchers
and in static maps to open access mapping systems available to the
public. Some of the most relevant existing tools are summarised below
and presented in Table~\ref{tbl-tools}.

@@

The Climate Action Plan scenario is loosely based on the Irish
Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which contains policies for action to achieve a 51\%
reduction in overall GHG emissions by 2030, on a path towards net-zero
by 2050. For transport, this includes 500,000 extra walking, cycling and
public transport trips per day by 2030. In terms of car travel, the
target is to ``Increas{[}e{]} the proportion of kilometers driven by
passenger electric cars to between 40 and 45\% by 2030, in addition to a
reduction of 10\% in kilometers driven by the remaining internal
combustion engine cars.'' This equates to a 5.5 to 6\% reduction in
total car km driven.

@@

@@

  \bibliography{bibliography.bib,cruse.bib}

>>>>

Transport systems dominated by heavy and powerful private cars are
inefficient, dangerous, and unhealthy. Cars are a major and growing
source of greenhouse gas emissions causing climate change
\citep{winkler2023}, a leading cause of premature death and injury due
to road traffic collisions \citep{globals2018} , and a cause of disease
due to physical inactivity and air, noise and (from tyres) microplastic
pollution \citep{mattsson2023} \citep{welch2023} \citep{cavallaro2024}.
Transport is responsible for 23\% of global greenhouse gas (GHG)
emissions, 70\% of which are from road transport, with passenger cars
accounting for nearly half of transport emissions (around 10\% of global
emissions) \citep{jaramillo2022}. The transport system encourages,
enables and in some cases enforces unsustainable lifestyles. Services
that are only accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

@@

This paper describes an approach to strategic cycle network planning
tool development that is open, scalable, and evidence-based. We label
the approach Cycle Route Uptake and Scenario Estimation (CRUSE) and
present a case study of its application in Ireland, the results of which
are publicly available at \href{https://cruse.bike/}{cruse.bike} in the
CRUSE Tool for Ireland, the landing page of which is illustrated in
Figure~\ref{fig-landing}. In Section~\ref{sec-methods}, we outline the
methods used to generate the evidence presented in the CRUSE Tool. In
Section~\ref{sec-results}, we present the results of the CRUSE Tool,
including estimates of current cycling levels and future cycling
potential at the national, regional and local levels. In
Section~\ref{sec-discussion}, we discuss limitations and possible future
improvements to the approach, and the implications of the results for
Ireland before making concluding remarks in
Section~\ref{sec-conclusions}.

@@

Tools for estimating geographical distribution of cycling potential have
advanced substantially in recent years. Developed by researchers and
developers in academic, public and private sectors, they have evolved
from simple area-based static models tailored to specific regions to
more complex and (in some cases) more generalizable tools. These tools
can be considered as a specific application of the concept of planning
support systems (PSS) , which are designed to help planners and
decision-makers to make more informed decisions \citep{geertman2009}.
Outputs include maps with results available at several levels of
analysis with levels of availability ranging from being only available
to researchers and in static maps to open access mapping systems
available to the public.

Previous studies have applied the concept of PSS to specific modes,
including walking \citep{bencekri2024}, public transport
\citep{barmentlo2012} and cycling \citep{bencekri2023}. While such prior
work provides insight into the potential for approaches to transport
planning that are both data-driven and participatory, the focus of this
paper is on tools that focus attention on the \emph{geographic
distribution} of cycling potential and which answer the question ``where
to build''. Not all papers reviewed in this section classify themselves
as, or even mention, PSS, but all of them can be considered as PSS. This
paper is informed by the thinking underlying PSS, including the idea
that the resulting tools are more effective if they are modular and can
inter-link with other tools and processes in the planning process: PSS
can be conceived as a ``toolbox with separate instruments that can work
and talk to each other effectively (as in the open systems concept) and
that will be selectively applied in varying configurations to support a
particular phase in the planning process'' \citep{geertman2002}. We
share this conception of PSS, emphasizing the benefits of a focus on a
particular mode (cycling) and a particular phase in the planning process
(strategic network development).

Some of the most relevant existing tools are summarised below and
presented in Table~\ref{tbl-tools}.

@@

The Climate Action Plan scenario models the transport emissions
reductions targeted in the Irish Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which aims for a 51\% cut in overall greenhouse gas
emissions by 2030 as part of the pathway to net-zero emissions by 2050.
For transport, this includes 500,000 extra walking, cycling and public
transport trips per day by 2030. In terms of car travel, the target is
to ``Increas{[}e{]} the proportion of kilometers driven by passenger
electric cars to between 40 and 45\% by 2030, in addition to a reduction
of 10\% in kilometers driven by the remaining internal combustion engine
cars.'' This equates to a 5.5 to 6\% reduction in total car km driven.

@@

CRUSE is a ``great tool for getting started with project outline
documents''. It's an ``ideal tool for generating a strategic rationale
for estimating demand and building the business case in the absence of
cycle count data'' (Declan Keenan, Senior Transport Planner working on
active travel, personal communication). ``It's a good representation of
what's going on''

``I like to know that the data on the county maps is from POWSCAR data
which is gold standard''.

@@

  \bibliography{bibliography.bib,cruse.bib,references.bib}
